### 2022-12-11 latex/beale_best_fits.tex
First Commit

\begin{tabular}{llrrr}
\toprule
Algorithm &                               Solution &  Iterations &  Function Evaluations &  Function Value \\
\midrule
      dfp &  2.9999999997259486,0.4999999999232928 &           8 &                    10 &    1.378756e-20 \\
     bfgs &   3.000000000000004,0.5000000000000007 &          15 &                    18 &    8.125884e-30 \\
    lbfgs &  3.0000000003576988,0.5000000000625404 &          14 &                    50 &    3.625854e-20 \\
      lma & 2.9999997171743287,0.49999992960730744 &          12 &                    13 &    1.280491e-14 \\
\bottomrule
\end{tabular}