### 2022-12-11 latex/matyas_best_fits.tex

\begin{tabular}{llrrr}
\toprule
Algorithm &                                        Solution &  Iterations &  Function Evaluations &  Function Value \\
\midrule
      dfp &   1.7926757117875478e-11,1.7863785568107216e-11 &           7 &                    10 &    1.281062e-23 \\
     bfgs & -4.0285184756435043e-19,-4.0615726129211233e-19 &           7 &                    10 &    6.547689e-39 \\
    lbfgs &   -3.837208328860697e-15,4.6629367034256575e-15 &           4 &                    16 &    1.806993e-29 \\
      lma &   3.5389027698051246e-06,3.5389027799876804e-06 &          14 &                    15 &    5.009533e-13 \\
\bottomrule
\end{tabular}