### 2022-12-11 latex/matyas_function_values.tex

\begin{tabular}{lrrrr}
\toprule
Algorithm &      Minimal &      Maximum &         Mean &       Median \\
\midrule
      dfp & 1.281062e-23 & 1.159689e-09 & 6.085427e-11 & 5.222055e-12 \\
     bfgs & 6.547689e-39 & 6.314078e-18 & 2.829248e-19 & 4.146166e-22 \\
    lbfgs & 1.806993e-29 & 4.906348e-10 & 1.455184e-11 & 5.802267e-14 \\
      lma & 5.009533e-13 & 2.403290e-09 & 1.680342e-10 & 2.454947e-11 \\
\bottomrule
\end{tabular}