### 2022-12-11 latex/rastrigin2d_best_fits.tex

\begin{tabular}{llrrr}
\toprule
Algorithm &                                       Solution &  Iterations &  Function Evaluations &  Function Value \\
\midrule
      dfp & -1.3945825916973396e-12,1.3762564161053067e-12 &           6 &                    11 &             0.0 \\
     bfgs &  4.190758074827946e-12,-2.1862133845293774e-11 &           8 &                    49 &             0.0 \\
    lbfgs &  8.784224615873606e-12,-3.6460273411768743e-11 &           5 &                   133 &             0.0 \\
      lma &  2.6660763338975985e-10,1.5611908894664897e-09 &           4 &                     5 &             0.0 \\
\bottomrule
\end{tabular}