### 2022-12-11 latex/rastrigin2d_function_values.tex

\begin{tabular}{lrrrr}
\toprule
Algorithm &  Minimal &      Maximum &         Mean &       Median \\
\midrule
      dfp &      0.0 & 1.492140e-13 & 1.186606e-14 & 0.000000e+00 \\
     bfgs &      0.0 & 0.000000e+00 & 0.000000e+00 & 0.000000e+00 \\
    lbfgs &      0.0 & 1.989918e+00 & 1.989918e-02 & 0.000000e+00 \\
      lma &      0.0 & 3.552714e-15 & 1.882938e-15 & 3.552714e-15 \\
\bottomrule
\end{tabular}