### 2022-12-11 latex/rastrigin30d_function_values.tex

\begin{tabular}{lrrrr}
\toprule
Algorithm &  Minimal &      Maximum &         Mean &  Median \\
\midrule
      dfp &      0.0 & 1.648459e-12 & 5.002221e-14 &     0.0 \\
     bfgs &      0.0 & 0.000000e+00 & 0.000000e+00 &     0.0 \\
    lbfgs &      0.0 & 0.000000e+00 & 0.000000e+00 &     0.0 \\
      lma &      0.0 & 5.684342e-14 & 1.477929e-14 &     0.0 \\
\bottomrule
\end{tabular}