### 2022-12-11 latex/rastrigin4d_best_fits.tex

\begin{tabular}{llrrr}
\toprule
Algorithm &                                           Solution &  Iterations &  Function Evaluations &  Function Value \\
\midrule
      dfp & 7.824099927923849e-11,-2.1366231163644037e-10,-... &          10 &                    21 &             0.0 \\
     bfgs & 5.405272390841542e-12,-2.7239120065655507e-11,-... &          11 &                    61 &             0.0 \\
    lbfgs & 1.471102762075013e-10,1.9145855829354901e-10,8.... &           5 &                   133 &             0.0 \\
      lma & 2.6660763338975985e-10,1.5611908894664897e-09,-... &           4 &                     5 &             0.0 \\
\bottomrule
\end{tabular}