### 2022-12-11 latex/rosenbrock2d_best_fits.tex

\begin{tabular}{llrrr}
\toprule
Algorithm &                              Solution &  Iterations &  Function Evaluations &  Function Value \\
\midrule
      dfp & 0.9999999999008999,0.9999999997978861 &          24 &                    31 &    1.135257e-20 \\
     bfgs &                1.0,0.9999999999999999 &          30 &                    43 &    1.232595e-30 \\
    lbfgs & 0.9999999994699225,0.9999999989299816 &          27 &                    76 &    2.907107e-19 \\
      lma & 0.9999999979097256,0.9999999958107717 &          16 &                    18 &    4.376780e-18 \\
\bottomrule
\end{tabular}