### 2022-12-11 latex/rosenbrock2d_function_values.tex

\begin{tabular}{lrrrr}
\toprule
Algorithm &      Minimal &      Maximum &         Mean &       Median \\
\midrule
      dfp & 1.135257e-20 & 1.330822e+00 & 3.883331e-02 & 1.008079e-13 \\
     bfgs & 1.232595e-30 & 5.420407e-20 & 1.520616e-21 & 3.568745e-24 \\
    lbfgs & 2.907107e-19 & 6.027957e-11 & 1.010571e-12 & 1.481586e-15 \\
      lma & 4.376780e-18 & 1.308671e-10 & 7.473415e-12 & 8.945853e-13 \\
\bottomrule
\end{tabular}