### 2022-12-11 latex/rosenbrock30d_best_fits.tex

\begin{tabular}{llrrr}
\toprule
Algorithm &                                           Solution &  Iterations &  Function Evaluations &  Function Value \\
\midrule
      dfp & 1.000000001426901,1.0000000035108567,1.00000000... &        2767 &                  2790 &    2.303328e-14 \\
     bfgs & 1.0000000000000022,1.0,0.9999999999999846,0.999... &         207 &                   231 &    2.745279e-23 \\
    lbfgs & 1.0000000005707574,1.0000000006791032,1.0000000... &         160 &                   389 &    4.773145e-16 \\
      lma & 1.0,0.9999999999999999,0.9999999999999999,0.999... &          53 &                   131 &    1.018105e-16 \\
\bottomrule
\end{tabular}