### 2022-12-11 latex/rosenbrock30d_function_values.tex

\begin{tabular}{lrrrr}
\toprule
Algorithm &      Minimal &   Maximum &      Mean &       Median \\
\midrule
      dfp & 2.303328e-14 & 29.257503 & 13.357326 & 1.307731e+01 \\
     bfgs & 2.745279e-23 &  3.986624 &  0.797325 & 6.316386e-22 \\
    lbfgs & 4.773145e-16 &  3.986624 &  0.199331 & 4.316852e-14 \\
      lma & 1.018105e-16 &  3.986624 &  0.478395 & 2.761710e-13 \\
\bottomrule
\end{tabular}