### 2022-12-11 latex/rosenbrock4d_best_fits.tex

\begin{tabular}{llrrr}
\toprule
Algorithm &                                           Solution &  Iterations &  Function Evaluations &  Function Value \\
\midrule
      dfp & 0.9999999966340929,0.9999999907960823,0.9999999... &         292 &                   301 &    3.610838e-15 \\
     bfgs & 1.0000000000000053,1.0000000000000098,1.0000000... &          29 &                    35 &    6.042675e-28 \\
    lbfgs & 0.999999999931242,0.999999999845467,0.999999999... &          45 &                   134 &    1.861496e-19 \\
      lma & 0.9999999999421001,0.9999999998839636,0.9999999... &          18 &                    20 &    7.114597e-20 \\
\bottomrule
\end{tabular}