### 2022-12-11 latex/rosenbrock4d_function_values.tex

\begin{tabular}{lrrrr}
\toprule
Algorithm &      Minimal &  Maximum &     Mean &       Median \\
\midrule
      dfp & 3.610838e-15 & 3.707671 & 0.601430 & 1.338231e-03 \\
     bfgs & 6.042675e-28 & 3.701429 & 0.370143 & 1.679398e-23 \\
    lbfgs & 1.861496e-19 & 3.701429 & 0.407157 & 2.624568e-15 \\
      lma & 7.114597e-20 & 3.701429 & 0.333129 & 1.322880e-13 \\
\bottomrule
\end{tabular}