### 2022-12-13 latex/ackley_best_solutions.tex

\begin{table}
\centering
\caption{Detailed Solutions For Ackley Function}
\label{detailedsolutions:ackley}
\begin{tabular}{lrrrr}
\toprule
 Coord. &  $S_{1}$ &  $S_{2}$ &  $S_{3}$ &  $S_{4}$ \\
\midrule
$x_{1}$ &    -0.00 &     0.00 &     0.00 &     0.00 \\
$x_{2}$ &    -0.00 &     0.00 &     0.00 &     0.00 \\
\bottomrule
\end{tabular}
\end{table}