### 2022-12-13 latex/ackley_convergence.tex

\begin{table}
\centering
\caption{Convergence Report For Ackley Function}
\label{convergence:ackley}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Good &  Poor &  Diver. &  Total \\
\midrule
  dfp &   100 &     0 &       0 &    100 \\
 bfgs &   100 &     0 &       0 &    100 \\
lbfgs &   100 &     0 &       0 &    100 \\
  lma &   100 &     0 &       0 &    100 \\
\bottomrule
\end{tabular}
\end{table}