### 2022-12-13 latex/beale_convergence.tex

\begin{table}
\centering
\caption{Convergence Report For Beale Function}
\label{convergence:beale}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Good &  Poor &  Diver. &  Total \\
\midrule
  dfp &    53 &    25 &      22 &    100 \\
 bfgs &    75 &     0 &      25 &    100 \\
lbfgs &    80 &     0 &      20 &    100 \\
  lma &    42 &    37 &      21 &    100 \\
\bottomrule
\end{tabular}
\end{table}