### 2022-12-13 latex/beale_function_values.tex

\begin{table}
\centering
\caption{Statistical Information about function values For Beale Function}
\label{function_values:beale}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Min &   Max &  Mean &  Median \\
\midrule
  dfp & 0.00 &  3.84 &  0.27 &    0.00 \\
 bfgs & 0.00 &  0.45 &  0.11 &    0.00 \\
lbfgs & 0.00 &  7.31 &  0.36 &    0.00 \\
  lma & 0.00 & 14.20 &  5.35 &    0.45 \\
\bottomrule
\end{tabular}
\end{table}