### 2022-12-13 latex/booth_best_fits.tex

\begin{table}
\centering
\caption{Best Fits For Booth Function}
\label{solutions:booth}
\begin{tabular}{llrrr}
\toprule
 Alg. &    Sol. &  Iter. &  F. Eval &  F. Value \\
\midrule
  dfp & $S_{1}$ &      4 &        6 &      0.00 \\
 bfgs & $S_{2}$ &      2 &        5 &      0.00 \\
lbfgs & $S_{3}$ &      5 &       21 &      0.00 \\
  lma & $S_{4}$ &     10 &       11 &      0.00 \\
\bottomrule
\end{tabular}
\end{table}