### 2022-12-13 latex/rastrigin2d_best_fits.tex

\begin{table}
\centering
\caption{Best Fits For Rastrigin Function}
\label{solutions:rastrigin2d}
\begin{tabular}{llrrr}
\toprule
 Alg. &    Sol. &  Iter. &  F. Eval &  F. Value \\
\midrule
  dfp & $S_{1}$ &      6 &       11 &      0.00 \\
 bfgs & $S_{2}$ &      8 &       49 &      0.00 \\
lbfgs & $S_{3}$ &      5 &      133 &      0.00 \\
  lma & $S_{4}$ &      4 &        5 &      0.00 \\
\bottomrule
\end{tabular}
\end{table}