### 2022-12-13 latex/rastrigin2d_function_values.tex

\begin{table}
\centering
\caption{Statistical Information about function values For Rastrigin Function}
\label{function_values:rastrigin2d}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Min &  Max &  Mean &  Median \\
\midrule
  dfp & 0.00 & 0.00 &  0.00 &    0.00 \\
 bfgs & 0.00 & 0.00 &  0.00 &    0.00 \\
lbfgs & 0.00 & 1.99 &  0.02 &    0.00 \\
  lma & 0.00 & 0.00 &  0.00 &    0.00 \\
\bottomrule
\end{tabular}
\end{table}