### 2022-12-13 latex/rosenbrock30d_best_fits.tex

\begin{table}
\centering
\caption{Best Fits For Rosenbrock Function}
\label{solutions:rosenbrock30d}
\begin{tabular}{llrrr}
\toprule
 Alg. &    Sol. &  Iter. &  F. Eval &  F. Value \\
\midrule
  dfp & $S_{1}$ &   2967 &     2990 &      0.00 \\
 bfgs & $S_{2}$ &    236 &      252 &      0.00 \\
lbfgs & $S_{3}$ &    164 &      406 &      0.00 \\
  lma & $S_{4}$ &     53 &      131 &      0.00 \\
\bottomrule
\end{tabular}
\end{table}