### 2022-12-13 latex/rosenbrock30d_convergence.tex

\begin{table}
\centering
\caption{Convergence Report For Rosenbrock Function}
\label{convergence:rosenbrock30d}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Good &  Poor &  Diver. &  Total \\
\midrule
  dfp &     1 &    99 &       0 &    100 \\
 bfgs &    80 &    20 &       0 &    100 \\
lbfgs &    95 &     5 &       0 &    100 \\
  lma &    88 &    12 &       0 &    100 \\
\bottomrule
\end{tabular}
\end{table}