### 2022-12-13 latex/rosenbrock30d_function_values.tex

\begin{table}
\centering
\caption{Statistical Information about function values For Rosenbrock Function}
\label{function_values:rosenbrock30d}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Min &   Max &  Mean &  Median \\
\midrule
  dfp & 0.00 & 29.25 & 13.10 &   12.58 \\
 bfgs & 0.00 &  3.99 &  0.80 &    0.00 \\
lbfgs & 0.00 &  3.99 &  0.20 &    0.00 \\
  lma & 0.00 &  3.99 &  0.48 &    0.00 \\
\bottomrule
\end{tabular}
\end{table}