### 2022-12-14 latex/ackley_best_fits.tex
Rewritting latex files

The best solutions of all algorithms for Ackley Function, obtained using the minimal
distance of the solution, are shown in Table~\ref{solutions:ackley}:

\begin{table}[H]
\centering
\caption{Best Fits For Ackley Function}
\label{solutions:ackley}
\begin{tabular}{llrrr}
\toprule
 Alg. &    Sol. &  Iter. &  F. Eval &  F. Value \\
\midrule
  dfp & $S_{1}$ &     36 &      108 &      0.00 \\
 bfgs & $S_{2}$ &     42 &      129 &      0.00 \\
lbfgs & $S_{3}$ &     37 &      389 &      0.00 \\
  lma & $S_{4}$ &     38 &      320 &      0.00 \\
\bottomrule
\end{tabular}
\end{table}

The Ackley function has a narrow space around the optimal solution and,
altough specially LMA is of rapid convergence, it iterated 38 times to find
the optimal value, probably because it walked around the solution
before reach it. All algorithms were very slow to converge and the most expensive
are LMA and LBFGS.

Altough the column F. Value points to the the same value, this is a rounded value
and the real values are os order $10^{-22} \approx 10^{-14}$.