### 2022-12-14 latex/ackley_function_values.tex

The minimal. maximal, mean and median values of the solutions are shown in Table~\ref{function_values:ackley}:

\begin{table}[H]
\centering
\caption{Statistical Information about function values For Ackley Function}
\label{function_values:ackley}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Min &  Max &  Mean &  Median \\
\midrule
  dfp & 0.00 & 0.00 &  0.00 &    0.00 \\
 bfgs & 0.00 & 0.00 &  0.00 &    0.00 \\
lbfgs & 0.00 & 0.00 &  0.00 &    0.00 \\
  lma & 0.00 & 0.00 &  0.00 &    0.00 \\
\bottomrule
\end{tabular}
\end{table}

The values again are rounded. The matter is the mdian value: it informs us the result that
we are like to obtain when running any algorithm to solve for minimum. As it seems, for all
algorithms is expected to find the minimun of Ackley Function.