### 2022-12-14 latex/beale_best_fits.tex

The best solutions of all algorithms for Beale Function, obtained using the minimal
distance of the solution, are shown in Table~\ref{solutions:beale}:

\begin{table}[H]
\centering
\caption{Best Fits For Beale Function}
\label{solutions:beale}
\begin{tabular}{llrrr}
\toprule
 Alg. &    Sol. &  Iter. &  F. Eval &  F. Value \\
\midrule
  dfp & $S_{1}$ &      8 &       10 &      0.00 \\
 bfgs & $S_{2}$ &     15 &       18 &      0.00 \\
lbfgs & $S_{3}$ &     14 &       50 &      0.00 \\
  lma & $S_{4}$ &     12 &       13 &      0.00 \\
\bottomrule
\end{tabular}
\end{table}

The Beale function has a very large search space around the optimal solution and
it had a reasonable number of iterations for each algorithm. The DFP method, superseeded by BFGS,
is surprisingly rapid to find the solution. The LMA, known to be fast, had a reasonable performance.
BFGS/LBFGS performed as expected.

Altough the column F. Value points to the the same value, this is a rounded value
and the real values are os order $10^{-22} \approx 10^{-14}$.