### 2022-12-14 latex/beale_best_solutions.tex

The best solutions of all algorithms for Beale Function, indicated as
$S_{1}$, $S_{2}$, $S_{3}$ and $S_{4}$ in Table~\ref{solutions:beale}, are shown
in Table~\ref{detailedsolutions:beale}:

\begin{table}[H]
\centering
\caption{Detailed Solutions For Beale Function}
\label{detailedsolutions:beale}
\begin{tabular}{lrrrr}
\toprule
 Coord. &  $S_{1}$ &  $S_{2}$ &  $S_{3}$ &  $S_{4}$ \\
\midrule
$x_{1}$ &     3.00 &     3.00 &     3.00 &     3.00 \\
$x_{2}$ &     0.50 &     0.50 &     0.50 &     0.50 \\
\bottomrule
\end{tabular}
\end{table}

The optimal solution for Beale Function is $[3, 0.5]$ and the rounded values
of Table~\ref{detailedsolutions:beale} points to the same value. The error of
these values are of order $10^{-13}$ os less.