### 2022-12-14 latex/beale_function_values.tex

The minimal. maximal, mean and median values of the solutions are shown in Table~\ref{function_values:beale}:

\begin{table}[H]
\centering
\caption{Statistical Information about function values For Beale Function}
\label{function_values:beale}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Min &   Max &  Mean &  Median \\
\midrule
  dfp & 0.00 &  3.84 &  0.27 &    0.00 \\
 bfgs & 0.00 &  0.45 &  0.11 &    0.00 \\
lbfgs & 0.00 &  7.31 &  0.36 &    0.00 \\
  lma & 0.00 & 14.20 &  5.35 &    0.45 \\
\bottomrule
\end{tabular}
\end{table}

LMA again had the worst performance. It is the only algorithm that, altiugh has some solutions in
poor convergence, that has a median greater than zero. LBFGS and DFP seems quite radical in terms
of convergence, while having a median of zero its mean values are greater than zero an its maximal
value are noticeable. The larger mean value of LMA reinforces the notion of divergence in this case
and signifies that the algorithm don't fit very well for BEale Fnction minimum.