### 2022-12-19 latex/matyas_best_fits.tex

The best solutions of all algorithms for Booth Function, obtained using the minimal
distance of the solution, are shown in Table~\ref{solutions:matyas}:

\begin{table}[H]
\centering
\caption{Best Fits For Matyas Function}
\label{solutions:matyas}
\begin{tabular}{llrrr}
\toprule
 Alg. &    Sol. &  Iter. &  F. Eval &  F. Value \\
\midrule
  dfp & $S_{1}$ &      7 &       10 &      0.00 \\
 bfgs & $S_{2}$ &      7 &       10 &      0.00 \\
lbfgs & $S_{3}$ &      4 &       16 &      0.00 \\
  lma & $S_{4}$ &     14 &       15 &      0.00 \\
\bottomrule
\end{tabular}
\end{table}

The algorithms DFP, BFGS and LBFGS have comparable performances, while LMA
is the most expensive of all.

Although the column F. Value points to the the same value, this is a rounded value
and the real values are os order $10^{-22} \approx 10^{-14}$.