### 2022-12-19 latex/matyas_best_solutions.tex

The best solutions of all algorithms for Matyas Function, indicated as
$S_{1}$, $S_{2}$, $S_{3}$ and $S_{4}$ in Table~\ref{solutions:matyas}, are shown
in Table~\ref{detailedsolutions:matyas}:

\begin{table}[H]
\centering
\caption{Detailed Solutions For Matyas Function}
\label{detailedsolutions:matyas}
\begin{tabular}{lrrrr}
\toprule
 Coord. &  $S_{1}$ &  $S_{2}$ &  $S_{3}$ &  $S_{4}$ \\
\midrule
$x_{1}$ &     0.00 &    -0.00 &    -0.00 &     0.00 \\
$x_{2}$ &     0.00 &    -0.00 &     0.00 &     0.00 \\
\bottomrule
\end{tabular}
\end{table}

The optimal solution for Matyas Function is $\left[0, 0\right]$ and the rounded values
of Table~\ref{detailedsolutions:matyas} points to the same value. The error of
these values are of order $10^{-13}$ os less.