### 2022-12-19 latex/matyas_convergence.tex

The convergence report for the Matyas Function is shown in Table~\ref{convergence:matyas}:

\begin{table}[H]
\centering
\caption{Convergence Report For Matyas Function}
\label{convergence:matyas}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Good &  Poor &  Diver. &  Total \\
\midrule
  dfp &   100 &     0 &       0 &    100 \\
 bfgs &   100 &     0 &       0 &    100 \\
lbfgs &   100 &     0 &       0 &    100 \\
  lma &   100 &     0 &       0 &    100 \\
\bottomrule
\end{tabular}
\end{table}

All algorithms converged 100 times, using 100 different initial solutions,
for a search space of $\left[-10, 10\right]$ . In this search space,
Matyas Function has a well defined valley that has the solution as its
center, as it is shown in Section~\ref{isolinesmatyas2D}.