### 2022-12-19 latex/rastrigin2d_convergence.tex

The convergence report for the Rastrigin Function in 2 dimensions is shown in Table~\ref{convergence:rastrigin2d}:

\begin{table}[H]
\centering
\caption{Convergence Report For Rastrigin Function}
\label{convergence:rastrigin2d}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Good &  Poor &  Diver. &  Total \\
\midrule
  dfp &   100 &     0 &       0 &    100 \\
 bfgs &   100 &     0 &       0 &    100 \\
lbfgs &    99 &     1 &       0 &    100 \\
  lma &   100 &     0 &       0 &    100 \\
\bottomrule
\end{tabular}
\end{table}

It is quite sure that all algorithms has good convergence for any algorithm for a
constrained search space of $\left[-0.15, 0.15\right]$. This search space is around
the optimal solution to avoid wrong convergence to a local minimum. In this search space,
Rastrigin Function has a well defined valley that has the solution as its center, as is shown in
Section~\ref{isolinesrastrigin2D}.