### 2022-12-19 latex/rastrigin2d_function_values.tex

The minimal. maximal, mean and median values of the solutions are shown in Table~\ref{function_values:rastrigin2d}:

\begin{table}[H]
\centering
\caption{Statistical Information about function values For Rastrigin Function}
\label{function_values:rastrigin2d}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Min &  Max &  Mean &  Median \\
\midrule
  dfp & 0.00 & 0.00 &  0.00 &    0.00 \\
 bfgs & 0.00 & 0.00 &  0.00 &    0.00 \\
lbfgs & 0.00 & 1.99 &  0.02 &    0.00 \\
  lma & 0.00 & 0.00 &  0.00 &    0.00 \\
\bottomrule
\end{tabular}
\end{table}

The values again are rounded. The median informs us that, for all
algorithms is expected to find the minimun of Rastrigin Function.
Although the mean of LBFGS algorithm is greater than zero, it is expected
to converge mora than diverge in the function constrained search space.