### 2022-12-19 latex/rastrigin30d_best_fits.tex

The best solutions of all algorithms for Rastrigin Function, obtained using the minimal
distance of the solution, are shown in Table~\ref{solutions:rastrigin30d}:

\begin{table}[H]
\centering
\caption{Best Fits For Rastrigin Function}
\label{solutions:rastrigin30d}
\begin{tabular}{llrrr}
\toprule
 Alg. &    Sol. &  Iter. &  F. Eval &  F. Value \\
\midrule
  dfp & $S_{1}$ &      8 &       14 &      0.00 \\
 bfgs & $S_{2}$ &      9 &       54 &      0.00 \\
lbfgs & $S_{3}$ &      6 &      135 &      0.00 \\
  lma & $S_{4}$ &      4 &        5 &      0.00 \\
\bottomrule
\end{tabular}
\end{table}

LMA is the less expensive algorithm to solve Rastrigin in 30 dimensions, while LBFGS
is the most expensive of all. Although it has one more iteration than LMA, it evaluates
the objective function twice than BFGS, the second most expensive algorithm.

Although the column F. Value points to the the same value, this is a rounded value
and the real values are os order $10^{-22} \approx 10^{-14}$.