### 2022-12-19 latex/rastrigin4d_function_values.tex

The minimal. maximal, mean and median values of the solutions are shown in Table~\ref{function_values:rastrigin4d}:

\begin{table}[H]
\centering
\caption{Statistical Information about function values For Rastrigin Function}
\label{function_values:rastrigin4d}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Min &  Max &  Mean &  Median \\
\midrule
  dfp & 0.00 & 0.00 &  0.00 &    0.00 \\
 bfgs & 0.00 & 0.00 &  0.00 &    0.00 \\
lbfgs & 0.00 & 0.00 &  0.00 &    0.00 \\
  lma & 0.00 & 0.00 &  0.00 &    0.00 \\
\bottomrule
\end{tabular}
\end{table}

The values again are rounded. The median informs us that, for all
algorithms is expected to find the minimun of Rastrigin Function.