### 2022-12-19 latex/rosenbrock2d_best_fits.tex

The best solutions of all algorithms for Rosenbrock Function, obtained using the minimal
distance of the solution, are shown in Table~\ref{solutions:rosenbrock2d}:

\begin{table}[H]
\centering
\caption{Best Fits For Rosenbrock Function}
\label{solutions:rosenbrock2d}
\begin{tabular}{llrrr}
\toprule
 Alg. &    Sol. &  Iter. &  F. Eval &  F. Value \\
\midrule
  dfp & $S_{1}$ &     24 &       31 &      0.00 \\
 bfgs & $S_{2}$ &     30 &       43 &      0.00 \\
lbfgs & $S_{3}$ &     27 &       76 &      0.00 \\
  lma & $S_{4}$ &     16 &       18 &      0.00 \\
\bottomrule
\end{tabular}
\end{table}

LMA is the less expensive algorithm to solve Rosenbrock Function in 2 dimensions, while LBFGS
is the most expensive of all. Although it has comparable iterations with DFP nd BFGS, it evaluates
the objective function twice as much of BFGS, the second most expensive algorithm.

Although the column F. Value points to the the same value, this is a rounded value
and the real values are os order $10^{-22} \approx 10^{-14}$.