### 2022-12-19 latex/rosenbrock2d_best_solutions.tex

The best solutions of all algorithms for Rosenbrock Function, indicated as
$S_{1}$, $S_{2}$, $S_{3}$ and $S_{4}$ in Table~\ref{solutions:rosenbrock2d}, are shown
in Table~\ref{detailedsolutions:rosenbrock2d}:

\begin{table}[H]
\centering
\caption{Detailed Solutions For Rosenbrock Function}
\label{detailedsolutions:rosenbrock2d}
\begin{tabular}{lrrrr}
\toprule
 Coord. &  $S_{1}$ &  $S_{2}$ &  $S_{3}$ &  $S_{4}$ \\
\midrule
$x_{1}$ &     1.00 &     1.00 &     1.00 &     1.00 \\
$x_{2}$ &     1.00 &     1.00 &     1.00 &     1.00 \\
\bottomrule
\end{tabular}
\end{table}

The optimal point of Rosenbrock function in 2 dimensions is correctly pointed to $\left[1, 1\right]$. The Table~\ref{detailedsolutions:rosenbrock2d}
shows the same solution to all algorithms, because the values are so small and were rounded to fit on the Table. The differences of
the optimal point are of order $10^{-14}$ to $10^{-22}$, for the LMA, DFP, LBFGS and BFGS algorithms, respectively.