### 2022-12-19 latex/rosenbrock2d_convergence.tex

The convergence report for the Rosenbrock Function in 30 dimensions is shown in Table~\ref{convergence:rosenbrock2d}:

\begin{table}[H]
\centering
\caption{Convergence Report For Rosenbrock Function}
\label{convergence:rosenbrock2d}
\begin{tabular}{lrrrr}
\toprule
 Alg. &  Good &  Poor &  Diver. &  Total \\
\midrule
  dfp &    74 &    26 &       0 &    100 \\
 bfgs &   100 &     0 &       0 &    100 \\
lbfgs &   100 &     0 &       0 &    100 \\
  lma &   100 &     0 &       0 &    100 \\
\bottomrule
\end{tabular}
\end{table}

Except for DFP, Rosenbrock has good convergence for any algorithm for a constrained search space
of $\left[-1.5, 1.5\right]$. Although the Rosenbrock Function has only the global minimum in a
search space of $\left[-10, 10\right]$, it has a large banana shaped plateau near the optimal
solution, as it is shown in Section~\ref{isolinesrosenbrock2D}. The search space constraint serves
only to limit the number of iterations of each algorithm. The large search space tends to make the algorithm
walks too much to find the solution and tends to exceed the maximum iterations defined in the evaluation
criteria shown in Section~\ref{sec:introduction}.